---
title: "Exploratory Analysis of CEO Salary Data"
author: "Felipe Campos, Josh Wilson and Stanimir Vichev"
date: "1/24/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("ggplot2")
```

# Introduction

The focus of this study is to try to determine whether a company's performance is related to its CEO salary. To accomplish that, data gathered from a selection of different companies will be used and the whole treatment and analysis of the data will be documented in this study along with relevant findings. The conclusions taken from these findings will then be discussed.

It is important for the purpose of this study, however, to initially discuss the design of this analysis and the concepts used to operationalize it. The most relevant concept used throughout this report is company performance, given that it is what we will be measuring against CEO data. We can use a conceptual definition that a successful company is one that has an increased market value and is able to consistently deliver results. Therefore, to measure a company's performance this concept will be operationalized in this study by measuring that company's market value as well as its profitability. Therefore, a company could be considered as presenting successful performance if it delivered profits in accordance to its market value. 

# Possible inclusion of discussion about operationalization of salary and company performance

# Handling of Variable Coding Issues and Missing Values

	Our careful examination of the data revealed that some variables, such as mktval and profits, had missing values, and that certain entries appeared to be erroneous. An initial check revealed no NA values, but looking at the mktval variable we could see some -1 values. 
```{r}
CEO[!complete.cases(CEO),]
CEO[which(CEO$mktval == -1),]
```

As a company’s market value cannot be negative, we reasoned that -1 indicates missing data. As the five rows that has missing mktval also had -1 for profits, we reasoned that the same coding principle was used for that variable as well. While profits can be negative, it looks like -1 was exclusively used for missing data.
```{r}
CEO[which(CEO$profits == -1),]
```

The missing values were all in our key variables, so we couldn’t proceed with these rows in our analysis. Therefore, we set the -1 values to NA and removed the five rows from the set.
```{r}
CEO$mktval[CEO$mktval == -1] <- NA
CEO$profits[CEO$profits == -1] <- NA
CEO_complete <- CEO[complete.cases(CEO),]
```

During our analysis, we tried to determine if any instances contained erroneous data that could impact our results. Checking whether any of the CEOs had comten less than ceoten (meaning that they worked in the company fewer years than they were CEO there), we found one row with such wrong values. 
```{r}
CEO_complete[CEO_complete$ceoten > CEO_complete$comten, ]
```

We also saw two CEOs that had gone to grad school, but not to college, which seems highly unlikely to be true. Furthermore, the two rows seem to be almost exactly the same, apart from the age. Similar to the above, it looks like these values could have been entered twice incorrectly.
```{r}
CEO_complete[CEO_complete$grad > CEO_complete$college, ]
```

We decided to remove these three wrong rows, so that they didn’t influence any conclusions we drew from the data set.
```{r}
ceoten_comten_error = which(CEO_complete$ceoten > CEO_complete$comten)
CEO_complete <- CEO_complete[-ceoten_comten_error,]
grad_college_error = which(CEO_complete$grad > CEO_complete$college)
CEO_complete <- CEO_complete[-grad_college_error,]
```

# Data and Univariate Analysis

We begin our study by loading the data gathered from the companies being studied and analyzing the main variables available.

```{r data load and summary}
load("ceo_w203.RData")
names(CEO)
```

In the dataset used in this report, we have the following data measured from selected companies:

- **CEO Salary**, which is the compensation for the CEO in the year 1990, measured in thousands of dollar;
- **CEO Age**, measuring the age of the CEO in years;
- **CEO College Education**, indicating whether the CEO completed college education, being 1 for yes and 0 for no;
- **CEO Graduate Education**, indicating whether the CEO completed graduate level education, being 1 for yes and 0 for no;
- **CEO Total Time Within Company**, which measures the time the CEO has been an employee of that company in years;
- **CEO Time Within Company as CEO**, measuring the time the CEO has held that position in the company in years;
- **Company Profitability**, which measures the profitability for that company in the year of 1990 in millions of dollars;
- **Company Market Value**, indicating the market value for the company in the year of 1990 in millions of dollars;

Each of theses variables will be now thoroughly analyzed and discussed independently.

## CEO Salary

We begin by analyzing the histogram of the CEO Salary variable.

```{r salary histogram}
salary_hist <- ggplot(data = CEO, aes(x = salary))
salary_hist + geom_histogram(bins = 40)
```

The histogram of the dependent variable salary uncovers two important features about the salary sample. First, the distribution has a positive skew, with many salaries bunched into bins between 500 and 1000 and a long tail towards the right. The mean being greater than the median demonstrates the effect of the right tail. Second, there are two or more possible outliers. The positive skew is illustrated through the fact that the leftmost values of the graph are only approximately 1.5 standard deviations away from the mean, while the higher values can exceed this. From this evidence, we believe that distribution for salary is non-normal and may require transformation, especially if we are to use models that assume normally distributed variables.  We also note that it is not possible to have a salary less than 0. Moreover, we understand that outliers often tell very interesting stories in their own right, and we have labeled the clear outlier as the "rich ceo". At this moment, we cannot discard the "rich ceo" because there is no evidence that this is an error.

## CEO Age

```{r}
age_hist <- ggplot(data = CEO_complete, aes(x = age))
age_hist + geom_histogram(bins = 60)
```

The distribution of age in the sample looks fairly normal, except for the distinct peak at the center. From the kurtosis calculation, we might say this sample has more in common with a logistic distribution than a normal distribution. Of course, this is only a sample of 185, and we might find that kurtosis will approach 0 as the sample grows. However, it is likely that the lack of heavy tails is an important feature of CEO age, because we see the sparseness in CEOs below 30 and over 75. We could easily come up with an ad hoc story suggesting that CEOs below 30 don't have the requisite experience and CEOs above 75 cannot handle the rigors of the job. Moreover, the CEOs who are older than 75 may be CEOs in name only, with much of the business operations carried out by others. <<This might be verified by having very low salaries and/or very low market value companies.>> Another view would be to label points above 75 and below 30 as outliers, but this would be hasty without further information. We note that the mean is fairly high at 55 years old and the standard deviation is also rather large at 9.5. 

## CEO College Education


## CEO Graduation Education


## CEO Total Time Within Company


## CEO Time Within Company as CEO


## Company Profitability


## Company Market Value


# Data Treatment



# Analysis of Bivariate Relationships (uses cleaned data and transformations)

Salary and market value
```{r}
salary_v_mktval <- ggplot(data = CEO_complete, aes(x = log10(CEO_complete$mktval), 
                                           y = log10(CEO_complete$salary))) +
  geom_point(alpha=.5) + geom_smooth() +
  xlab("Company Market Value in USD (Millions) on Log Scale") +
  ylab("CEO Salary in USD (1000s) on Log Scale") +
  theme(legend.position = c(1,1), legend.justification= c(1,1), legend.background = element_blank(),
        plot.title = element_text(hjust = .5)) +
  labs(title = "CEO Salary vs. Company Market Value on Log-Log Scale")
  
salary_v_mktval
cor(log10(CEO_complete$mktval),log10(CEO_complete$salary)) # 0.481491
```

Salary and education
```{r}
# The two lines below will be moved up to the "coding issues" section.
CEO_complete$ed <- factor(CEO_complete$college + CEO_complete$grad)
levels(CEO_complete$ed) <- c("No College", "College", "Grad School")

salary_v_ed <- ggplot(data = CEO_complete, aes(x = factor(ed), y = salary))
salary_v_ed + geom_boxplot() + 
  stat_summary(fun.y="mean", geom="point", shape = 23, size =3, fill = "red") +
  xlab("Level of Education") + ylab("Salary in USD (1000s)") +
  labs(title="CEO Salary vs. Level of Education") +
  theme(legend.position = c(1,1), legend.justification= c(1,1), legend.background = element_blank(),
  plot.title = element_text(hjust = .5))
```

Salary vs. Experience (Years as CEO)
```{r}
salary_v_ceoten <- ggplot(data = CEO_complete, aes(x = CEO_complete$ceoten, y = log10(CEO_complete$salary))) +
  geom_point(position = position_jitter(width = 0.25, height = 0),alpha = 0.5) + geom_smooth() +
  xlab("Years") + ylab("Salary in USD (1000s) on Log Scale") + labs(title = "Salary vs. Years as CEO") +   scale_x_continuous(breaks = seq(0,35,5)) + 
  theme(plot.title = element_text(hjust = .5))
salary_v_ceoten
cor(CEO_complete$salary, CEO_complete$ceoten) # 0.1429477
cor(log10(CEO_complete$salary), CEO_complete$ceoten) # 0.114728
```

# Looking at secondary variables relationships




Bivariate analysis of Market Value vs. Profit
```{r}
mktval_v_profits <- ggplot(data = CEO_complete, aes(x = log10(CEO_complete$mktval), 
                                           y = log10(CEO_complete$profits + 500))) +
  geom_point(position = position_jitter(width = 0, height = 0.25), alpha=.5) + geom_smooth() +
  xlab("Company Market Value in USD (Millions) on Log Scale") +
  ylab("Company Profits in USD (Millions) on Log Scale") +
  theme(legend.position = c(1,1), legend.justification= c(1,1), legend.background = element_blank(),
        plot.title = element_text(hjust = .5)) +
  labs(title = "Company Market Value vs. Company Profits on Log-Log Scale")
  
mktval_v_profits
cor(log10(CEO_complete$mktval),log10(CEO_complete$profits + 500)) # 0.6988081
```

# Possibly including the relationship of salary in the above graph
```{r}
mktval_v_profits_v2 <- ggplot(data = CEO_complete, aes(x = log10(CEO_complete$mktval), 
                                           y = log10(CEO_complete$profits + 500),
                                           colour = log10(CEO_complete$salary))) +
  geom_point(position = position_jitter(width = 0.25, height = 0.25), alpha=.5) + geom_smooth() +
  xlab("Company Market Value in USD (Millions) on Log Scale") +
  ylab("Company Profits in USD (Millions) on Log Scale") +
  theme(legend.position = c(1,0), legend.justification= c(1,0), legend.background = element_blank(),
        plot.title = element_text(hjust = .5)) +
  labs(title = "Company Market Value vs. Company Profits on Log-Log Scale", colour ="Log Salary")
  
mktval_v_profits_v2
```


Tenure as CEO vs. Age with Company Tenure

# Potential Confounding Effects in Bivariate Relationships
```{r}
CeoTen_v_tenure <- ggplot(data = CEO_complete, aes(x = CEO_complete$ceoten, 
                                           y = CEO_complete$age,
                                           colour = CEO_complete$comten)) +
  geom_point(position = position_jitter(width = 0.0, height = 0.25), alpha=.5) + geom_smooth() +
  xlab("Years as CEO") +
  ylab("Years on Earth") +
  theme(legend.position = c(1,0), legend.justification= c(1,0), legend.background = element_blank(),
        plot.title = element_text(hjust = .5)) +
  labs(title = "CEO Tenure vs. Age", colour ="Years at Co") +
  scale_color_gradient(low = "blue", high = "white") + scale_x_continuous(breaks = seq(0,35,5))
  
CeoTen_v_tenure
```



# Conclusion

