---
title: "Exploratory Analysis of CEO Salary Data"
author: "Felipe Campos, Josh Wilson and Stanimir Vichev"
date: "1/24/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("ggplot2")
```

# Introduction

The focus of this study is to try to determine whether a company's performance is related to its CEO salary. To accomplish that, data gathered from a selection of different companies will be used and the whole treatment and analysis of the data will be documented in this study along with relevant findings. The conclusions taken from these findings will then be discussed.

It is important for the purpose of this study, however, to initially discuss the design of this analysis and the concepts used to operationalize it. The most relevant concept used throughout this report is company performance, given that it is what we will be measuring against CEO data. We can use a conceptual definition that a successful company is one that has an increased market value and is able to consistently deliver results. Therefore, to measure a company's performance this concept will be operationalized in this study by measuring that company's market value as well as its profitability. Therefore, a company could be considered as presenting successful performance if it delivered profits in accordance to its market value. 

# Data and Univariate Analysis

We begin our study by loading the data gathered from the companies being studied and analyzing the main variables available.

```{r data load and summary}
load("ceo_w203.RData")
names(CEO)
```

In the dataset used in this report, we have the following data measured from selected companies:

- **CEO Salary**, which is the compensation for the CEO in the year 1990, measured in thousands of dollar;
- **CEO Age**, measuring the age of the CEO in years;
- **CEO College Education**, indicating whether the CEO completed college education, being 1 for yes and 0 for no;
- **CEO Graduate Education**, indicating whether the CEO completed graduate level education, being 1 for yes and 0 for no;
- **CEO Total Time Within Company**, which measures the time the CEO has been an employee of that company in years;
- **CEO Time Within Company as CEO**, measuring the time the CEO has held that position in the company in years;
- **Company Profitability**, which measures the profitability for that company in the year of 1990 in millions of dollars;
- **Company Market Value**, indicating the market value for the company in the year of 1990 in millions of dollars;

Each of theses variables will be now thoroughly analyzed and discussed independently.

## CEO Salary

We begin by analyzing the histogram of the CEO Salary variable.

```{r salary histogram}
salary_hist <- ggplot(data = CEO, aes(x = salary))
salary_hist + geom_histogram(bins = 40)
```

The histogram of the dependent variable salary uncovers two important features about the salary sample. First, the distribution has a positive skew, with many salaries bunched into bins between 500 and 1000 and a long tail towards the right. The mean being greater than the median demonstrates the effect of the right tail. Second, there are two or more possible outliers. The positive skew is illustrated through the fact that the leftmost values of the graph are only approximately 1.5 standard deviations away from the mean, while the higher values can exceed this. From this evidence, we believe that distribution for salary is non-normal and may require transformation, especially if we are to use models that assume normally distributed variables.  We also note that it is not possible to have a salary less than 0. Moreover, we understand that outliers often tell very interesting stories in their own right, and we have labeled the clear outlier as the "rich ceo". At this moment, we cannot discard the "rich ceo" because there is no evidence that this is an error.

## CEO Age

```{r}
age_hist <- ggplot(data = CEO_complete, aes(x = age))
age_hist + geom_histogram(bins = 60)
```

The distribution of age in the sample looks fairly normal, except for the distinct peak at the center. From the kurtosis calculation, we might say this sample has more in common with a logistic distribution than a normal distribution. Of course, this is only a sample of 185, and we might find that kurtosis will approach 0 as the sample grows. However, it is likely that the lack of heavy tails is an important feature of CEO age, because we see the sparseness in CEOs below 30 and over 75. We could easily come up with an ad hoc story suggesting that CEOs below 30 don't have the requisite experience and CEOs above 75 cannot handle the rigors of the job. Moreover, the CEOs who are older than 75 may be CEOs in name only, with much of the business operations carried out by others. <<This might be verified by having very low salaries and/or very low market value companies.>> Another view would be to label points above 75 and below 30 as outliers, but this would be hasty without further information. We note that the mean is fairly high at 55 years old and the standard deviation is also rather large at 9.5. 

## CEO College Education


## CEO Graduation Education


## CEO Total Time Within Company


## CEO Time Within Company as CEO


## Company Profitability


## Company Market Value


# Data Treatment



# Analysis of Bivariate Relationships



# Potential Confounding Effects in Bivariate Relationships



# Conclusion

