---
title: "Exploratory Analysis of CEO Salary Data"
author: "Felipe Campos, Stanimir Vichev, Josh Wilson"
date: "2/2/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("ggplot2")
library("gridExtra")
library("knitr")
library("car")
library(e1071)
```

# Introduction

The purpose of this exploratory data analysis is to determine whether there is an association between the salary of a chief executive officer (CEO) and the performance of the CEO’s company. To that effect, we analyzed a cross section of data gathered from a selection of different companies for the year 1990. In our analysis, we will demonstrate that there is a strong linear association between a CEO’s salary and the performance of a company, after the appropriate variable transformations have been implemented. Moreover, we will show that a CEO’s experience may also play a role in how much she receives in salary.

Initially we will discuss how the constructs were operationalized in the study. This will lead us into an investigation of each variable’s sample distribution. From the univariate analysis, we will discuss oddities we found in the data and the necessary steps we took to create a clean data set. Then, we explore the bivariate relationships within the data, including a section on potential confounding effects. Lastly, we reiterate our key findings in the conclusion as well as some apparent limitations to the study. 

# Operationalization of Constructs in Analysis  

Before we can begin an analysis of the data, we have to determine the key concepts and how they were operationalized. If there is too much flexibility in how an operationalized variable can be interpreted, then the study design will be flawed and the results must be taken with a grain of salt. In our study, the salient yet undefined concept is company performance. We can define a successful company as one that has increased its market value over time and is able to consistently deliver results. Therefore, we can infer from this intuitive definition that measuring a company's market value should provide a perspective into company performance. (We note that are many different ways to calculate market value, and we cannot be sure how this term is defined in the data set.) However, many investors interpret a successful company as that which delivers high profitability versus share size. Therefore, “company performance” could also be reasonably measured by profits. 

A more meaningful measure of company performance might come from simple feature engineering. Profits as a standalone measure is fairly meaningless without taking into account the size of the company. In which case, a ratio of profits to the market value might provide unforeseen insight into the relationship of company performance and a CEO’s salary.

# Data and Univariate Analysis

We begin our study by loading the data gathered from the companies being studied and analyzing the main variables available.

```{r data load and summary}
load("ceo_w203.RData")
names(CEO)
```

In the dataset used in this report, we have the following data measured from selected companies:

- **CEO Salary**, which is the compensation for the CEO in the year 1990, measured in thousands of dollar;
- **CEO Age**, measuring the age of the CEO in years;
- **CEO College Education**, indicating whether the CEO completed college education, being 1 for yes and 0 for no;
- **CEO Graduate Education**, indicating whether the CEO completed graduate level education, being 1 for yes and 0 for no;
- **CEO Total Time Within Company**, which measures the time the CEO has been an employee of that company in years;
- **CEO Time Within Company as CEO**, measuring the time the CEO has held that position in the company in years;
- **Company Profitability**, which measures the profitability for that company in the year of 1990 in millions of dollars;
- **Company Market Value**, indicating the market value for the company in the year of 1990 in millions of dollars;

Each of theses variables will be now thoroughly analyzed and discussed independently.

## CEO Salary

We begin by analyzing the histogram of the CEO Salary variable.

```{r salary histogram}
salary_hist <- ggplot(data = CEO, aes(x = salary)) + geom_histogram(fill = "white", colour = "black", bins = 40) + xlab("Salary in USD (1000s)") + ylab("Frequency") + 
  labs(title="Distribution of CEO Salary") +
  theme(legend.position = c(1,1), legend.justification= c(1,1), legend.background = element_blank(),
  plot.title = element_text(hjust = .5)) + 
  geom_vline(aes(xintercept=mean(salary), colour = "mean"), linetype = "dashed") + 
  geom_vline(aes(xintercept=median(salary), colour = "median"), linetype = "dashed") + 
  scale_color_manual(name = "Statistics", values = c(median = "blue", mean = "red")) +
  scale_x_continuous(breaks = c(0, 1000, 2000, 3000, 4000, 5000)) + scale_y_continuous(breaks = c(0, 5, 10, 15, 20, 25, 30))
salary_hist
```

The histogram of the dependent variable salary uncovers two important features about the salary sample. First, the distribution has a positive skew, with many salaries bunched into bins between 500 and 1000 and a long tail towards the right. The mean being greater than the median demonstrates the effect of the right tail. Second, there are two or more possible outliers. The positive skew is illustrated through the fact that the leftmost values of the graph are only approximately 1.5 standard deviations away from the mean, while the higher values can exceed this. From this evidence, we believe that distribution for salary is non-normal and may require transformation, especially if we are to use models that assume normally distributed variables.  We also note that it is not possible to have a salary less than zero and the presence of only non-negative salaries confirms this intuition. Moreover, we understand that outliers often tell very interesting stories in their own right, and we have labeled the clear outlier as the "rich ceo". At this moment, we cannot discard the "rich ceo" because there is no evidence that this is an error.


## Company Market Value
```{r}
mktval_hist <- ggplot(data = CEO, aes(x=CEO$mktval)) + 
  geom_histogram(fill = "white", colour = "black", bins = 40) + xlab("Market Value in USD (Millions)") + ylab("Frequency") + 
  labs(title="Distribution of Market Value") +
  theme(legend.position = c(1,1), legend.justification= c(1,1), legend.background = element_blank(),
  plot.title = element_text(hjust = .5)) + 
  geom_vline(aes(xintercept=mean(CEO$mktval, na.rm = T), colour = "mean"), linetype = "dashed") + 
  geom_vline(aes(xintercept=median(CEO$mktval, na.rm = T), colour = "median"), linetype = "dashed") + 
  scale_color_manual(name = "Statistics", values = c(median = "blue", mean = "red")) +
  scale_x_continuous(breaks = c(0, 5000, 10000, 15000, 20000, 25000, 30000, 35000, 40000, 45000)) + 
  scale_y_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60))
mktval_hist

sd(CEO$mktval, na.rm = T)
kurtosis(CEO$mktval, na.rm=T)
```
The market value histogram appears in line with our intuition about the stock market; most companies have lower market values around 7m (USD), with only a few going above 2.5b (USD). Furthermore, the histogram shows that companies with very large market values (above $4b) are rare and can be considered as potential outliers, as a host of values are greater than three standard deviations from the mean. Therefore, these “rich companies” can have an interesting effect on our analysis and should be considered carefully. The presence of these outliers creates a heavy right tail and the extreme peak towards the lower end of the market value spectrum suggests that a log transformation will be needed to accommodate issues of potential heteroskedasticity.

While investigating the market value variable, we found that there were certain companies with a market value of negative one. As it is impossible for a company to have negative market value and there aren’t any other negative values for this variable, we reasoned that negative one signifies missing data. Comparing the histograms of the market value variable with and without the negative one rows, we see a small shift in the distribution of the variable near zero, together with a small increase of the mean. As we removed a few values close to zero, this small shift is to be expected.


## Company Profitability
```{r}
profits_hist <- ggplot(data = CEO, aes(x = profits)) + 
  geom_histogram(fill = "white", colour = "black", bins = 40) + 
  xlab("Profits in USD (Millions)") + ylab("Frequency") +
  labs(title="Distribution of Company Profits in USD (Millions)") +
  theme(legend.position = c(1,1), legend.justification= c(1,1), legend.background = element_blank(),
  plot.title = element_text(hjust = .5)) + 
  geom_vline(aes(xintercept=mean(CEO$profits, na.rm = T), colour = "mean"), linetype = "dashed") + 
  geom_vline(aes(xintercept=median(CEO$profits, na.rm = T), colour = "median"), linetype = "dashed") + 
  scale_color_manual(name = "Statistics", values = c(median = "blue", mean = "red")) +
  scale_x_continuous(breaks = c(-500, 0, 500, 1000, 1500, 2000, 2500, 3000)) + 
  scale_y_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60))
profits_hist

mean(CEO$profits, na.rm = T)
sd(CEO$profits, na.rm=T)
kurtosis(CEO$profits, na.rm=T)
```
The distribution of profits is extraordinarily peaked near the zero value. Many of the companies in the sample do not make profits over 50 million, and a handful of companies are even in the red (as far as $463 million!). The skew of the distribution is very positive, pulling the mean up above 200, despite the prevalence of many values near zero. The peakedness and skew are confirmed by a very high value for kurtosis, above 11. The spread for the distribution is also rather large, at nearly 400. Even with a large standard deviation, there are many values over three standard deviations away from the mean in the positive direction. These outliers and the general shape of the profit distribution demand a transformation.

In conducting a routine inspection of the profits data, we discovered that there were five observations that contained negative one. We were also able to confirm that these are the same five rows that have negative one for their market value. However, by examining the range of the profit variable, we notice that there are other negative values. In fact, these other values appear reasonable since, unlike with market value, companies can have negative profits for a specific year in which their costs exceed their revenues. Nonetheless, as only the aforementioned five rows had profits of negative one, we reason that negative one represents missing profit value as well. Comparing the profit histograms with and without these rows, we can see that the impact of excluding these observations has a marginal effect on the shape of the distribution as well as the descriptive statistics.

## CEO Age

```{r, warning = FALSE}
age_hist <- ggplot(data = CEO, aes(x = age)) +
  geom_histogram(fill = "white", colour = "black", bins = 40) + 
  xlab("Years") + ylab("Frequency") +
  labs(title="Distribution of CEO Age") +
  theme(legend.position = c(1,1), legend.justification= c(1,1), legend.background = element_blank(),
  plot.title = element_text(hjust = .5)) + 
  geom_vline(aes(xintercept=mean(age), colour = "mean"), linetype = "dashed") + 
  geom_vline(aes(xintercept=median(age), colour = "median"), linetype = "dashed") + 
  scale_color_manual(name = "Statistics", values = c(median = "blue", mean = "red")) +
  scale_x_continuous(breaks = seq(20,90,5))
age_hist

sd(CEO$age)
kurtosis(CEO$age)
```

The distribution of the age variable in the sample appears fairly symmetric and normal, as the mean and median are close together. Furthermore, the kurtosis of the variable signifies that age might be viewed more as a logistic distribution rather than a normal one. The high peak at the center suggests that there may be an optimal age range for becoming a CEO: between 50 and 65. However, the sparseness in CEOs below 30 and over 75 is an important feature of the distribution; these points are over two standard deviations away from the mean. An ad hoc story is readily available: it is reasonable to infer that people below 30 may lack the requisite experience to lead publicly-traded companies, while people above 75 may not be able to handle the rigors of such a demanding position. While the points above 75 and below 30 years may represent outliers, it would be unreasonable to exclude them. It is possible that the lack of observations in these buckets is an artifact of the sample.

## CEO Education
```{r}
CEO$college <- as.factor(CEO$college)
CEO$grad <- as.factor(CEO$grad)
levels(CEO$grad) <- c("No Grad", "Grad")
levels(CEO$college) <- c("No College", "College")
kable(table(CEO$college, CEO$grad), format = "markdown")
```
For categorical variables such as college and grad, we examined their frequency distribution using tables. Evident from the table is that less than 4% of the CEOs in the sample didn't graduate from college. It would seem that post-secondary education is a necessary attribute for a CEO; this is not a surprising attribute of the data since performance in school tends to be an indicator of work ethics and/or intelligence. Moreover, boards that choose CEOs tend to prefer credentialed individuals as a means of mitigating risk. Meanwhile, the split between college and a post-tertiary degree are much closer. Meanwhile, there is less evidence to suggest that a post-tertiary degree confers much additional advantage over a college degree in becoming a CEO. While more education would be considered beneficial within the general population, it could be the case that extensive work and leadership experience is more valuable when senior management positions are filled. That is, the position for CEO tends to attract or possibly require highly educated individuals.

Finally, there are two CEOs that went to grad school, but not to college, which seems highly unlikely to be true. It is possible to imagine situations where people who have not gone to college go on to get honorary degrees from universities for special achievements, thus falling in the above case. However, since we have no information on the way the data was gathered, we cannot assume such honorary degrees were included in the study. Furthermore, these two CEOs seem to be almost exactly the same, apart from the age, as seen in their records from the sample shown below: 

```{r}
CEO[c(16,51),]
```
We suspect that this might be a case of erroneous entry that was recorded twice, meaning that these entries should be excluded from the analysis. As the comparison between the dirty (including two entries) and the clean data shows, the omission of these two variables does not have a meaningful effect on the results drawn from the study.

## CEO Total Time Within Company
```{r, warning = FALSE, message = FALSE}
ten_ceo_hist <- ggplot(data = CEO) + geom_histogram(aes(x = comten, fill = "comten"), alpha = 0.3, position = "identity") +
  geom_histogram(aes(x = ceoten, fill = "ceoten"), alpha = 0.3, position = "identity") + 
  scale_fill_manual(values = c(comten = "red", ceoten = "green")) +
  theme(legend.position = c(1,1), legend.justification= c(1,1), legend.background = element_blank(),
        plot.title = element_text(hjust = .5)) +
  labs(title = "Distribution of Years as CEO and of Years at Company", fill = "Tenure Type") +
  xlab("Years") + ylab("Count") +
  geom_vline(aes(xintercept=mean(comten), colour = "mean comten"), linetype = "dashed") +
  geom_vline(aes(xintercept=mean(ceoten), colour = "mean ceoten"), linetype = "dashed") + 
  scale_color_manual(name = "Statistics", values = c(mean = "blue", mean = "red")) +
  scale_x_continuous(breaks = c(0,10,20,30,40,50,60)) + scale_y_continuous(breaks = seq(0,35,5))
ten_ceo_hist
```
The comten variable appears to be bimodal, with one peak near 33 years and another at just under 10 years. This suggest that there may be two populations within this variable. The mean tenure of a CEO is just over 20 years with a very large standard deviation of approximately 12. However, the kurtosis of -1 suggests that we might be dealing with more of a uniformly distributed sample; this is certainly contrary to visual inspection, which we believe is more reliable in this case. We note the presence of a small right tail that skews the distribution a little bit. There is one apparent outlier: a CEO that has commited an entire lifetime to a company. <<We can discuss this further to make sure we are all in agreement>>

The ceoten variable has a heavily, positively skewed distribution with most of the CEO's having tenure under 10 years. The mean and standard deviation are nearly equal at about 7 years. Given how bunched the data is around the mean, we believe the long right tail and some outliers are inflating the average. This is also evidenced by the median being less than the mean. Kurtosis is highly positive, agreeing with the peaked nature of the histogram. None of the CEOs in the sample have been CEO for longer than 40 years. 

We note that both of these variables are left bounded, meaning they cannot go below zero.

# Data Treatment

# Handling of Variable Coding Issues and Missing Values

Our careful examination of the data revealed that some variables, such as mktval and profits, had missing values, and that certain entries appeared to be erroneous. An initial check revealed no NA values, but looking at the mktval variable we could see some -1 values. 
```{r}
load("ceo_w203.RData")
CEO[!complete.cases(CEO),]
CEO[which(CEO$mktval == -1),]
```

As a company’s market value cannot be negative, we reasoned that -1 indicates missing data. As the five rows that has missing mktval also had -1 for profits, we reasoned that the same coding principle was used for that variable as well. While profits can be negative, it looks like -1 was exclusively used for missing data.
```{r}
CEO[which(CEO$profits == -1),]
```

The missing values were all in our key variables, so we couldn’t proceed with these rows in our analysis. Therefore, we set the -1 values to NA and removed the five rows from the set.
```{r}
CEO$mktval[CEO$mktval == -1] <- NA
CEO$profits[CEO$profits == -1] <- NA
CEO_complete <- CEO[complete.cases(CEO),]
```

During our analysis, we tried to determine if any instances contained erroneous data that could impact our results. Checking whether any of the CEOs had comten less than ceoten (meaning that they worked in the company fewer years than they were CEO there), we found one row with such wrong values. 
```{r}
CEO_complete[CEO_complete$ceoten > CEO_complete$comten, ]
```

We also saw two CEOs that had gone to grad school, but not to college, which seems highly unlikely to be true. Furthermore, the two rows seem to be almost exactly the same, apart from the age. Similar to the above, it looks like these values could have been entered twice incorrectly.
```{r}
CEO_complete[CEO_complete$grad > CEO_complete$college, ]
```

We decided to remove these three wrong rows, so that they didn’t influence any conclusions we drew from the data set.
```{r}
ceoten_comten_error = which(CEO_complete$ceoten > CEO_complete$comten)
CEO_complete <- CEO_complete[-ceoten_comten_error,]
grad_college_error = which(CEO_complete$grad > CEO_complete$college)
CEO_complete <- CEO_complete[-grad_college_error,]
```

The univariate analysis of the variables demonstrated that some key variables appear to have skewed distributions. Specifically, salary, profit, and market value all have positively skewed distributions, which is to be expected when dealing with income-related variables. In order to clearly visualize and to understand the relationship between these variables, we applied a log transformation. This nonlinear transformation should assist us in detecting any linear relationships, if they exist.

The log transformed histogram for salary appears to resemble a normal distributed variable more closely, while the same cannot be said of the transformed market value and salary variables. The market value histogram shows marked improvement because the very long tail has been substantially reduced. However, the distribution retained some of its positive skew and appears lognormally distributed. Similar to the market value transformation, the histogram of the profit variable is less skewed than before, however, there remains a massive peak that is cause for concern. Since logarithmic transformations cannot handle negative values, we added a constant greater than the minimum value to each instance of profit. Moreover, much of the data is very tightly bunched and this will cause some non-constant variance issues for modeling, especially considering the presence of some potential influential outliers. The log transformations are rather interpretable for the profits and market value, but if we were insistent on models with strict assumptions of normality, we could attempt a box-cox transformation.

```{r}
salary_hist_final <- ggplot(data = CEO_complete, aes(x = log10(salary))) + geom_histogram(fill = "white", colour = "black", bins = 40) + xlab("Log Salary in USD (1000s)") + ylab("Frequency") + 
  labs(title="Histogram of CEO Salary on Log Scale") +
  theme(legend.position = c(1,1), legend.justification= c(1,1), legend.background = element_blank(),
  plot.title = element_text(hjust = .5)) + 
  geom_vline(aes(xintercept=mean(log10(salary)), colour = "mean"), linetype = "dashed") + 
  geom_vline(aes(xintercept=median(log10(salary)), colour = "median"), linetype = "dashed") + 
  scale_color_manual(name = "Statistics", values = c(median = "blue", mean = "red")) #+
  #scale_x_continuous(breaks = c(0, 1000, 2000, 3000, 4000, 5000)) + scale_y_continuous(breaks = c(0, 5, 10, 15, 20, 25, 30))
salary_hist_final
```

Market Value 
```{r}
mktval_hist_final <- ggplot(data = CEO_complete, aes(x=log10(mktval))) + 
  geom_histogram(fill = "white", colour = "black", bins = 40) + xlab("Market Value in USD (Millions) Log Scale") + ylab("Frequency") + 
  labs(title="Histogram of Market Value on Log Scale") +
  theme(legend.position = c(1,1), legend.justification= c(1,1), legend.background = element_blank(),
  plot.title = element_text(hjust = .5)) + 
  geom_vline(aes(xintercept=mean(log10(CEO_complete$mktval)), colour = "mean"), linetype = "dashed") + 
  geom_vline(aes(xintercept=median(log10(CEO_complete$mktval)), colour = "median"), linetype = "dashed") + 
  scale_color_manual(name = "Statistics", values = c(median = "blue", mean = "red"))
mktval_hist_final

```

```{r}
profits_hist_final <- ggplot(data = CEO_complete, aes(x = log10(profits + 500))) + 
  geom_histogram(fill = "white", colour = "black", bins = 60) + 
  xlab("Log Profits in USD (Millions)") + ylab("Frequency") +
  labs(title="Histogram of Company Profits in USD (Millions) Log Scale") +
  theme(legend.position = c(1,1), legend.justification= c(1,1), legend.background = element_blank(),
  plot.title = element_text(hjust = .5)) + 
  geom_vline(aes(xintercept=mean(log10(CEO_complete$profits + 500)), colour = "mean"), linetype = "dashed") + 
  geom_vline(aes(xintercept=median(log10(CEO_complete$profits + 500)), colour = "median"), linetype = "dashed") + 
  scale_color_manual(name = "Statistics", values = c(median = "blue", mean = "red")) #+
  #scale_x_continuous(breaks = c(-500, 0, 500, 1000, 1500, 2000, 2500, 3000)) + 
  #scale_y_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60))
profits_hist_final
```


# Analysis of Bivariate Relationships (uses cleaned data and transformations)

Salary and market value
```{r}
salary_v_mktval <- ggplot(data = CEO_complete, aes(x = log10(CEO_complete$mktval), 
                                           y = log10(CEO_complete$salary))) +
  geom_point(alpha=.5) + geom_smooth() +
  xlab("Company Market Value in USD (Millions) on Log Scale") +
  ylab("CEO Salary in USD (1000s) on Log Scale") +
  theme(legend.position = c(1,1), legend.justification= c(1,1), legend.background = element_blank(),
        plot.title = element_text(hjust = .5)) +
  labs(title = "CEO Salary vs. Company Market Value on Log-Log Scale")
  
salary_v_mktval
cor(log10(CEO_complete$mktval),log10(CEO_complete$salary)) # 0.481491
```

Salary and education
```{r}
# The two lines below will be moved up to the "coding issues" section.
CEO_complete$ed <- factor(CEO_complete$college + CEO_complete$grad)
levels(CEO_complete$ed) <- c("No College", "College", "Grad School")

salary_v_ed <- ggplot(data = CEO_complete, aes(x = factor(ed), y = salary))
salary_v_ed + geom_boxplot() + 
  stat_summary(fun.y="mean", geom="point", shape = 23, size =3, fill = "red") +
  xlab("Level of Education") + ylab("Salary in USD (1000s)") +
  labs(title="CEO Salary vs. Level of Education") +
  theme(legend.position = c(1,1), legend.justification= c(1,1), legend.background = element_blank(),
  plot.title = element_text(hjust = .5))
```

Salary vs. Experience (Years as CEO)
```{r}
salary_v_ceoten <- ggplot(data = CEO_complete, aes(x = CEO_complete$ceoten, y = log10(CEO_complete$salary))) +
  geom_point(position = position_jitter(width = 0.25, height = 0),alpha = 0.5) + geom_smooth() +
  xlab("Years") + ylab("Salary in USD (1000s) on Log Scale") + labs(title = "Salary vs. Years as CEO") +   scale_x_continuous(breaks = seq(0,35,5)) + 
  theme(plot.title = element_text(hjust = .5))
salary_v_ceoten
cor(CEO_complete$salary, CEO_complete$ceoten) # 0.1429477
cor(log10(CEO_complete$salary), CEO_complete$ceoten) # 0.114728
```

# Looking at secondary variables relationships




Bivariate analysis of Market Value vs. Profit
```{r}
mktval_v_profits <- ggplot(data = CEO_complete, aes(x = log10(CEO_complete$mktval), 
                                           y = log10(CEO_complete$profits + 500))) +
  geom_point(position = position_jitter(width = 0, height = 0.25), alpha=.5) + geom_smooth() +
  xlab("Company Market Value in USD (Millions) on Log Scale") +
  ylab("Company Profits in USD (Millions) on Log Scale") +
  theme(legend.position = c(1,1), legend.justification= c(1,1), legend.background = element_blank(),
        plot.title = element_text(hjust = .5)) +
  labs(title = "Company Market Value vs. Company Profits on Log-Log Scale")
  
mktval_v_profits
cor(log10(CEO_complete$mktval),log10(CEO_complete$profits + 500)) # 0.6988081
```

# Possibly including the relationship of salary in the above graph
```{r}
mktval_v_profits_v2 <- ggplot(data = CEO_complete, aes(x = log10(CEO_complete$mktval), 
                                           y = log10(CEO_complete$profits + 500),
                                           colour = log10(CEO_complete$salary))) +
  geom_point(position = position_jitter(width = 0.25, height = 0.25), alpha=.5) + geom_smooth() +
  xlab("Company Market Value in USD (Millions) on Log Scale") +
  ylab("Company Profits in USD (Millions) on Log Scale") +
  theme(legend.position = c(1,0), legend.justification= c(1,0), legend.background = element_blank(),
        plot.title = element_text(hjust = .5)) +
  labs(title = "Company Market Value vs. Company Profits on Log-Log Scale", colour ="Log Salary")
  
mktval_v_profits_v2
```


Tenure as CEO vs. Age with Company Tenure

# Potential Confounding Effects in Bivariate Relationships
```{r}
CeoTen_v_tenure <- ggplot(data = CEO_complete, aes(x = CEO_complete$ceoten, 
                                           y = CEO_complete$age,
                                           colour = CEO_complete$comten)) +
  geom_point(position = position_jitter(width = 0.0, height = 0.25), alpha=.5) + geom_smooth() +
  xlab("Years as CEO") +
  ylab("Years on Earth") +
  theme(legend.position = c(1,0), legend.justification= c(1,0), legend.background = element_blank(),
        plot.title = element_text(hjust = .5)) +
  labs(title = "CEO Tenure vs. Age", colour ="Years at Co") +
  scale_color_gradient(low = "blue", high = "white") + scale_x_continuous(breaks = seq(0,35,5))
  
CeoTen_v_tenure
```



# Conclusion

